### 2014-06-23 cyc-concepts/intro.tex
adds skeleton and authors

% Introduction :
% Why did I do the work?
% What were the central motivations and hypotheses?

\section{Introduction}
% The objectives of the work.

\emph{What will this work show?}

% The justification for the objectives.

\emph{Why is the work important?}

% Background

\emph{Who else has done what?}

\emph{How?}

\emph{What has this group previously done?}

% Guidance to the reader

\emph{What should the reader watch for in the paper?}

\emph{What are the interesting high points?}

\emph{What strategy did we use?}

% Summary/Conclusion

\emph{What should the reader expect as conclusion?}