### 2014-07-23 fundamentals/verification.tex
adds section for robustness of coding practices.

\section{Verification}
% Organize this section according to major topics
% give each topic a section heading in boldface.
% try to cover the major common points :
%
% problem design
% methods of measurement
% supporting models
% supporting data
% simulations run
% results

% Just write the section headings for each part and indicate what goes in that
% section with words :
%
% heading
% figures (with captions)
% schematics (with captions and footnotes)
% equations
% tables

% What does it mean?
% What did I actually test?
% What were the results?
% Did the work yield a new method?
% Did the work yield new knowledge?
% What measurements did I make?
% How were these measurements characterized?
% What methods were used?
% What were the results?
% How were the measurements made and characterized?

\subsection{Coding Practices}
% What have we done to ensure robustness ?
% Examples from Material class, Transactions, simulation IDs
\subsection{Testing Framework}
% Peer-review system (secretly a note on open source?)
\subsection{Benchmarks}
%Inpro benchmark?
%FCO code-to-code?