### 2014-11-02 fundamentals/conclusion.tex

\section{Conclusions}
% summarize the conclusions of the paper as a list of short phrases or
% sentences. Do not repeat the results section unless special emphasis is
% needed. The conclusion is just that, not a summary. It should add a new,
% higher level of analysis and should explicitly indicate the significance of
% the work.
% What does it all mean?
% What hypotheses were proven or disproven?
% What was learned?
% Why does it make a difference?

%Modularity encourages ecosystem development
%Enables cross-institutional collaboration
%Open paradigm allows community oversight, testing, verification
%Enables apples to apples comparisons

The \Cyclus nuclear fuel cycle framework presents a more generic and flexible 
alternative to existing fuel cycles simulators.

Where, previous nuclear fuel cycle simulators have had limited distribution, 
uncertain provenance, and restricted customizability, \Cyclus emphasizes an 
open access and development strategy.  Open development, with transparent 
publication of the source code, not only provides much-needed accessibility, 
but also enables transparency and community oversight.

The object-oriented \acrlong{ABM} simulation paradigm ensures more generic 
simulation capability. It allows \Cyclus to address common analyses in a more 
flexible fashion and enables analyses that are impossible with system dynamics 
codes.

Similarly, discrete, fidelity-agnostic object tracking facilitates simulations 
at every level of detail. Simulations relying on arbitrarily complex isotopic 
compositions are possible in \Cyclus, as are simulations not employing any 
physics at all. Indeed, agents of such varying fidelity can even exist in the 
same simulation. Researchers no longer need to reinvent the wheel in order to 
model a simulation focused on the aspects of the fuel cycle that their research 
focuses on.

Furthermore, when the capabilities within \Cyclus, \Cycamore, and the rest of 
the ecosystem are insufficient, adding custom functionality is simplified by a 
modular, plug-in architecture. A clean, modern \gls{API} simplifies 
customization and independent archetype development so that researchers can 
create models within their domain of expertise without modifying the core 
simulation kernel. Throughout the \Cyclus 
infrastructure, architecture choices have sought to enable cross-institutional 
collaboration and sustainable, community-driven development. The ecosystem 
of capabilities, already growing, may someday reflect the full diversity of use 
cases in the nuclear fuel cycle simulation domain.