### 2015-02-05 fundamentals/examples.tex
more details about example simulation.

\section{Demonstration}
% Organize this section according to major topics
% give each topic a section heading in boldface.
% try to cover the major common points :
%
% problem design
% methods of measurement
% supporting models
% supporting data
% simulations run
% results

% Just write the section headings for each part and indicate what goes in that
% section with words :
%
% heading
% figures (with captions)
% schematics (with captions and footnotes)
% equations
% tables

% What does it mean?
% What did I actually test?
% What were the results?
% Did the work yield a new method?
% Did the work yield new knowledge?
% What measurements did I make?
% How were these measurements characterized?
% What methods were used?
% What were the results?
% How were the measurements made and characterized?

The success of the \Cyclus simulator can be measured in many ways. 
The most compelling are demonstrations of the feasibility of the \Cyclus community-driven development 
model and of fundamental simulation capability. Nascent promising growth of the \Cyclus 
ecosystem at multiple institutions indicates that a fuel cycle simulator can 
advance in a community-driven development paradigm. Additionally, simulation 
results for both once-through and more complex recycling scenarios demonstrate that 
\Cyclus possesses the fundamental fuel cycle simulation capabilities to 
contribute to the field.

\subsection{Ecosystem}
% Cycamore library
% Cyder?

The \Cyclus community-driven software development model seeks to break the 
unproductive proliferation of overspecialized simulators. It instead leverages 
the collective expertise of fuel cycle researchers toward a single, more 
extensible, tool. Through the targeted
contributions of those researchers, an ecosystem of capabilities should emerge.
The \Cyclus `Ecosystem' is the collection of tools, calculation libraries, 
archetypes, data, and input files intended for use with the \Cyclus simulator. 
Members of the ecosystem include:
\begin{itemize}
\item the archetypes provided in the \Cycamore \cite{carlsen_cycamore_2014} 
repository
\item the archetypes created by researchers
\item isotopic composition data
\item historical facility deployment data
\item the \Cyclus \gls{GUI} tool Cyclist
\item fundamental analysis tools in the \Cyclus toolkit
\item tools for \Cyclus optimization, parallelization, and development
\end{itemize}
Taken together, these form an ecosystem of capabilities. Over time, this 
ecosystem will grow as archetype developers, kernel developers, and 
even users contribute capabilities developed for their own needs. Indeed, the 
long-term vision for the \Cyclus framework predicts an ever-expanding ecosystem 
of both general and specialized capability extensions.

Already, the ecosystem is growing. Early cross-institutional contributions to 
the ecosystem demonstrate a significant achievement by the \Cyclus framework 
and provide the basis for a community-driven development model.

\subsubsection{Supplementary Projects}

A number of projects and tools outside of the core simulation kernel have been 
developed to improve the scope and the diversity of the capabilities in the \Cyclus 
ecosystem. Table \ref{tab:coretools} lists the tools and projects developed 
under close integration with the \Cyclus kernel.  These tools are used to ease development
and simulation design (Cystub, Cycic, Ciclus), data visualization and analysis (Cyclist, Cyan),
and remote execution (Cloudlus).

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|r|X|r|}
\hline
\textbf{Name} & \textbf{Description} & \textbf{Citation} \\
\hline
Cycic &  Input control - embedded in Cyclist & \cite{flanagan_input_2013}\\
Cyclist & Interactive data exploration environment & \cite{livnat_cyclist_2014} \\
Ciclus & Continuous integration scripts for \Cyclus & \cite{scopatz_ciclus_2014}\\
Cycstub & Skeleton for clean slate module development & \cite{carlsen_cycstub_2014}\\
Cyan & \Cyclus analysis tool & \cite{carlsen_cyan_2014}\\
Cloudlus & Tools for running \Cyclus in a cloud environment & \cite{carlsen_cloudlus_2014} \\
\hline
\end{tabularx}
\caption{Many tools have been developed outside of the scope of the \Cyclus kernel for improved user, developer, and analyst experiences with \Cyclus.}
\label{tab:coretools}
\end{table}

\subsubsection{Archetype Contributions}

It is expected that the most common type of contribution to \Cyclus will be 
contributions of new archetypes. Researchers will be driven to create a new 
archetype when a need arises to represent a novel reactor type, an innovative 
reprocessing strategy, or a particular governmental or institutional behavior.  
The real-world utility of \Cyclus can in part be measured by the breadth and 
diversity of archetypes being developed in this way.

Early progress has been promising. Many archetypes external to the \Cycamore library (Table \ref{tab:cycamore})  
have been 
\cite{huff_streamblender_2014,huff_commodconverter_2014}
or are being 
\cite{flanagan_bright-lite_2014,skutnik_nuclear_2014,huff_mktdriveninst_2014} developed for contribution to the 
\Cyclus ecosystem. These archetypes provide the first examples of 
developer-contributed capabilities.  They add to the fundamental \Cycamore
archetypes by providing physics-based reactors, separation, fabrication, storage
facilities, and expanded institutional paradigms.

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|r|X|r|}
\hline
\textbf{Name} & \textbf{Description} & \textbf{Citation} \\
\hline
Bright-lite & A physics-based reactor archetype and fuel fabrication archetype & \cite{flanagan_bright-lite_2014} \\
Nuclear Fuel Inventory Model & A flexible, ORIGEN-based, reactor analysis module & \cite{skutnik_nuclear_2014} \\
CommodConverter & A simple commodity converting storage facility archetype  & \cite{huff_commodconverter_2014} \\
MktDrivenInst & An institution that controls deployment based on commodity availability & \cite{huff_mktdriveninst_2014} \\
SeparationsMatrix & A facility for elemental separations of used fuel & \cite{huff_streamblender_2014} \\
StreamBlender & A facility for fuel fabrication from reprocessed streams & \cite{huff_streamblender_2014} \\
\hline
\end{tabularx}
\caption{A diverse set of archetypes under development reflect the diverse 
needs of researchers at various institutions. These archetypes, contributed 
outside of the \Cyclus core and \Cycamore libraries are the first demonstration 
of community-driven development in a fuel cycle simulator.}
\label{tab:archetypes}
\end{table}

The existence and diversity of these contributed archetypes
illustrate the power and potential of the community-based development approach 
that \Cyclus has taken.

\subsection{Simulations}
%Inpro example (is this still running or did we deprecate it with 1.0?)

% MJG - INPRO should be tried again. @rwcarlsen has been running lots of
% simulations with the batch reactor, and its the second generation of my
% reactor models. We could easily use the same demand schedule and enrichment
% facility parameters and run it again.

This section will discuss the creation and results of a few simple fuel cycle
simulations to help illustrate the flexibility of \Cyclus. These simulations
are designed to illustrate Cyclus' capability.  Many simplifying assumptions
have been made with respect to material compositions, fuel transmutation, and
others. The three fuel cycles examined are:

\begin{enumerate}
    \item Once through (no recycle) 
    \item Modified open (1-pass recycle)
    \item closed (infinite-pass recycle)
\end{enumerate}

For each of these fuel cycles, a 1,100 month \Cyclus simulation with only a
single reactor is run in addition to a simulation with 10 reactors with
staggered refueling times.  As the number of staggered-cycle reactors
increases the system converges toward continuous material flow results.
\Cyclus flexibility allows this transition to be examined.  The facilities
used in these simulations are:

\begin{itemize}

    \item \textbf{Reactor} (\Class{cycamore::Reactor}): This is a reactor
        model that requests any number of input fuel types and uses static
        compositions to transmute them to after they are burnt and discharged
        from the core.  Due to the static spent fuel compositions spent fuel
        compositions become "incorrect" as the input fuel diverges from the
        fixed, expected input composition. The reactor was configured to model
        a light water reactor with a 3 batch core (20,000 kg HM per batch)
        operating on an 18 month cycle with a 2 month refuel time.  The
        reactor was also configured to take in either enriched UOX fuel or
        recycled MOX fuel - preferring the latter if available.

    \item \textbf{Recycled Fuel Fabrication} (\Class{cycamore::Fab}): This
        facility requests depleted uranium and separated fissile material and
        mixes them together into recycled MOX fuel that it offers to
        requesters.  The streams are mixed using a D-factors approach
        \TODO{add D-factors citation} in an attempt to match simple neutronics
        properties of an ideal target fuel composition.

    \item \textbf{Separations} (\Class{cycamore::Separations}): This facility
        takes in all kinds of spent fuel and separates it into plutonium and
        waste streams with some efficiency (0.99 was used for these
        simulations).  Up to 60,000 kg of Pu can be separated per month.

    \item \textbf{Repository} (\Class{cycamore::Sink}): This is an
        infinite-capacity facility that can take in all types of material
        including separations waste streams and spent reactor fuel.

    \item \textbf{UOX Fabrication} (\Class{cycamore::Source}): This facility
        provides enriched UOX fuel as requested with infinite capacity.

    \item \textbf{DU Source} (\Class{cycamore::Source}): This facility provides
        depleted uranium as requested with infinite capacity.

\end{itemize}

For the once through case, the \Class{Reactor} was set to offer all spent fuel
as waste.  For the modified open case, the \Class{Reactor} offerred spent UOX
into a spent fuel market where it is preferrentially sent to separations, but
spent MOX is still offerred as waste.  In the closed case, the \Class{Reactor}
offers all spent fuel into a spent fuel market. Because the separations
facility requests spent fuel with a higher preference than the repository, it
is sent there if possible.  All these preferences are easy to adjust and
\Cyclus will dynamically handle supply constraints and non-uniform preference
resolution.  It is notable that separations and recycle fuel fabrication
capacity is deployed identically in all simulations.  In the once-through
case, the recycling loop never acquires material, and so reactors always
receive fresh UOX fuel.  In the modified open case, the recycle loop only
acquires spent UOX.  The \gls{DRE} ensures everything operates smoothly in all
cases.

Cyclus' discrete materials make single-pass recycle easy to implement.  The
reactors keep track of fuel as discrete batches. A reactor remembers where it
received each batch from.  If a batch was received from the recycled fuel
fabrication facility, it does not offer it to separations and instead sends it
to the repository directly.  The material flows for the modified open case are
shown in Figure \ref{fig:flowmodopen}.

\begin{figure}[H]
\begin{center}
\includegraphics{./images/flow-mod-open-1.eps}
\end{center}
\caption{Modified open 1-pass \gls{MOX} recycle scenario material flows.}
\label{fig:flowmodopen}
\end{figure}

Changing the scenario from a modified open fuel cycle to a closed fuel cycle
requires only a one-word change in the input file regarding the output
commodity for the spent \gls{MOX} fuel of the \Class{Reactor}:

\begin{lstlisting}[language=diff]
     <fuel>         
       <incommodity>mox</incommodity>
  -    <outcommodity>waste</outcommodity>
  +    <outcommodity>spent_fuel</outcommodity>
       <inrecipe>mox_resh_fuel</inrecipe>
       <outrecipe>mox_spent_fuel</outrecipe>
     </fuel>
\end{lstlisting}

This results in the material flows in Figure \ref{fig:flowclosed}.  A
similarly trivial change was used to switch from the once through to a
modified open fuel cycle.  Note that the \Class{Reactor} always transmutes
fuel into fixed compositions, so nuclide-specific flows become
somewhat unrealistic in the fully closed cycle case.

\begin{figure}[H]
\begin{center}
\includegraphics{./images/flow-closed-1.eps}
\end{center}
\caption{Full \gls{MOX} recycle (multi-pass) fuel cycle material flows.}
\label{fig:flowclosed}
\end{figure}

Figure \ref{fig:puseries1} shows the full-system plutonium buildup for open
(no recycle), modified open, and closed (infinite-pass recycle) variations of
the one-reactor scenario described above. The figure was generated directly
from \Cyclus output data. After several batch cycles (near month 300) in the
modified open and closed cases, enough separated fissile material accumulates
in the fuel fabrication facility to generate a full recycled batch.  When this
batch is transmuted, more plutonium is burned than created.  This results in a
drop in the total fuel cycle system plutonium inventory.  This pattern repeats
roughly every 10 cycles (200 months) for the modified open case and every 9
cycles (180 months) for the closed case.  Because the modified open scenario
does not re-recycle material, it takes the fabrication facility 1 cycle longer
to accumulate a full batch of fissile material.

Because facilities are represented individually and transact discrete
materials as discrete events, realistic non-uniform patterns in facility
behavior that affect total system behavior are observed using \Cyclus.

\begin{figure}[H]
\begin{center}
\includegraphics{./images/puseries-1.eps}
\end{center}
\caption{System plutonium buildup with one reactor.}
\label{fig:puseries1}
\end{figure}

Figure \ref{fig:puseriesn} shows plutonium buildup for the 10-reactor
simulations of the open, modified open, and closed scenarios.  As the number
of reactors (with staggered refueling) increases, the behavior of the system
approaches a more steady average reminiscent of continuous material flow
models.

\begin{figure}[H]
\begin{center}
\includegraphics{./images/puseries-n.eps}
\end{center}
\caption{System plutonium buildup with staggered refueling for many reactors.}
\label{fig:puseriesn}
\end{figure}