### 2015-02-22 fundamentals/intro.tex
updates language between dynamic and static simulators. now all language points to dynamic as in time-stepping, facilities, etc.

% Introduction :

\section{Introduction}

% Why did we do the work?
% What were the central motivations and hypotheses?
% The objectives of the work?
% What will this work show the reader?
% Why is it important?
% What should the reader watch for in the paper?
% What are the interesting high points?
% What strategy did we use?
% What should the reader expect as conclusion?

As nuclear power expands, technical, economic, political, and environmental
analyses of nuclear fuel cycles by simulators increase in importance. The
merits of advanced nuclear technologies and fuel cycles are
shaped by myriad physical, nuclear, chemical, industrial, and political
factors. Nuclear fuel cycle simulators must therefore couple complex models of
nuclear process physics, facility deployment, and material routing. However,
current fuel cycle simulators rely on closed platforms and inflexible
architectures which exhibit three main failure modes. First, they discourage
targeted contribution and collaboration among experts. Next, they hobble
efforts to directly compare modeling methodologies. Finally, they
over-specialize, rendering most tools applicable to only a subset of desired
simulation fidelities, scales, and applications.

The cardinal purpose of a dyamic nuclear fuel cycle simulator is to calculate
the time- and facility-dependent mass flow through all or part the fuel cycle.
Dynamic nuclear fuel cycle analysis more realistically supports a range of
simulation goals than static analysis \cite{piet_dynamic_2011}. Historically,
dynamic nuclear fuel cycle simulators have calculated fuel cycle mass balances
and performance metrics derived from them using software ranging from ad hoc
spreadsheet-driven flow calculators to highly specialized system dynamics
modeling platforms. To date, current tools are often rigid rather than flexible,
fleet-based rather than discrete, or privately distributed rather than open
source. Each of these choices presents a challenge to modeling fidelity,
generality, efficiency, robustness, and scientific transparency.

The \Cyclus nuclear fuel cycle simulator framework and its modeling ecosystem
incorporate modern insights from simulation science and software architecture
to solve these problems.  These modern methods simultaneously enable more
efficient, accurate, robust, and validated analysis.  This next-generation fuel
cycle simulator is the result of design choices made to:

\begin{itemize}
\item support access to the tool by fuel cycle analysts and other users,
\item encourage developer extensions,
\item enable plug-and-play comparison of modeling methodologies,
\item and address a range of analysis types, levels of detail, and analyst sophistication.
\end{itemize}

\Cyclus, therefore, is a dynamic, agent-based model, which employs a modular
architecture, an open development process, discrete agents, discrete time, and
arbitrarily detailed isotopic resolution of materials. Experience in the broader
field of systems analysis indicates that agent-based modeling enables more
flexible simulation control, without loss of generality
\cite{macal_agent-based_2010}. Furthermore, openness allows cross-institutional
collaboration, increases software robustness \cite{cohen_modern_2010}, and
cultivates an ecosystem of modeling and calculation options. Comprised of
flexible, interchangeable models of fuel cycle components with varying scope,
depth and fidelity, this ecosystem allows users and developers to customise
\Cyclus to analyze the cases that are of interest to them, rather than those the
simulator was originally developed to address.  The fundamental concepts of the
\Cyclus nuclear fuel cycle simulator capture these modern insights so that novel
challenges in nuclear fuel cycle analysis can be better addressed.

\subsection{Background}

% Who else has done what?
% How did they do that?
% What did we do before this?
% What is new?

% Common goals of fuel cycle simulation/simulators (driving &d focus,
% approximating tech. impacts, etc.)
% Previous work (cite all the tools!)
% Limitations of previous simulators by others (fleet-based, closed
% development, inflexible infrastructure)
% Previous attempts by this group (GENIUS v1/v2)
% note NGFCS

Nuclear fuel cycle simulators drive \gls{RDD} by calculating `metrics',
quantitative measures of performance that can be compared among fuel cycle
options. The feasibility of the technology development and deployment
strategies which comprise a fuel cycle option, the dynamics of transitions
between fuel cycles, and many other measures of performance can be expressed in
terms of these metrics. For example, economic feasibility is often measured in
\gls{LCOE}, requiring calculation of lifetime
fuel costs and electricity generation, while environmental performance might be
measured by spent fuel volume, toxicity, or mined uranium.  A meta-analysis of
fuel cycle systems studies identified over two dozen unique quantitative
metrics spanning economics and cost, environmental sustainability and waste
management impacts, safety, security and nonproliferation, resource adequacy
and utilization, among others. With few exceptions, these metrics are derived
from mass balances calculated by a fuel cycle simulator.
\cite{flicker_evaluation_2014}.

However, methods for calculating those metrics vary among simulators. Some
model the system of facilities, economics, and materials in static equilibrium,
while other simulators capture the dynamics of the system.  Similarly, while
some simulators discretely model batches of material and individual facilities,
others aggregate facilities into fleets and materials into streams. Simulators
can model a single aspect of the fuel cycle in great detail while neglecting
others. For example, a simulator created for policy modeling might have
excellent capability in economics while capabailities for tracking transformations in
material isotopics and the effects of isotopics on technology performance are
neglected.  The \gls{CAFCA}\cite{guerin_impact_2009} simulator is problem-oriented
in this way, having elected to neglect isotopic resolution in favor of
integral effects.

Historically, domestic national laboratories have driven development and regulated
the use of their own tools: \gls{VISION}\cite{jacobson_verifiable_2010},
\gls{DANESS}\cite{van_den_durpel_daness_2009},
\gls{DYMOND}\cite{yacout_modeling_2005}, and
\gls{NFCSim}\cite{schneider_nfcsim:_2005,allan_guidance_2008}.
Internationally, other laboratories have created their own as well, such as
\gls{COSI}\cite{boucher_cosi_2005,boucher_cosi:_2006,meyer_new_2009,coquelet-pascal_comparison_2011}
and
\gls{ORION}\cite{worrall_scenario_2007}.
Outside the national laboratories,
researchers have created new nuclear fuel cycle simulation tools when existing
tools were not available or not sufficiently general to calculate their metrics
of interest.  With limited access to the national
laboratory tools and a need to customize them for research purposes,
universities and private industry researchers have ``reinvented the wheel'' by
developing tools of their own from scratch and tailored to their own needs.
Examples include \gls{CAFCA}\cite{guerin_benchmark_2009} and
\gls{DESAE}\cite{andrianova_desae_2008,mccarthy_benchmark_2012,allan_guidance_2008}.

\Cyclus emerged from a line of tools seeking to break this practice.  Its
precursor,
\gls{GENIUS} Version 1 \cite{dunzik-gougar_global_2007,jain_transitioning_2006},
originated within \gls{INL} and sought to provide generic regional capability.
Based on lessons learned from \gls{GENIUS} Version 1, the \gls{GENIUS} Version
2 \cite{oliver_studying_2009,huff_geniusv2_2009} simulator sought to provide more
generality and an extensible interface to facilitate collaboration.  The
\Cyclus project then improved upon the \gls{GENIUS} effort by
implementing increased modularity and encapsulation.  The result is a dynamic
simulator that treats both materials and facilities discretely, with an
architecture that permits multiple and variable levels of fidelity. The
simulator's agent-based framework for tracking the transformation and trade of
resources between autonomous regional and institutional entities with
customizable behavior and objectives is an innovation not pursued by any
existing fuel cycle simulator.

\subsection{Motivation}
% Gently reiterate the above need.
% There should be a hero narrative here
% There was a gap in the capabilities of simulators.
% \Cyclus, a hero, came in to fill that gap.

The \Cyclus paradigm enables targeted contribution and collaboration within the
nuclear fuel cycle analysis community to achieve two important goals: lower the
barrier for innvoative nuclear technologies to be included in fuel cycle analysis 
while improving the ability to compare simulations with and without those innovative
concepts.  This essential capability is absent in
previous simulators where user customization and extensibility were not design
objectives.  While the \emph{modular and open architecture} of
\Cyclus is necessary to meet these goals, it is not sufficient.  
\emph{Agent interchangeability} is required to facilitates direct comparison 
of alternative modeling methodologies and facility concepts.  Finally, \Cyclus is
applicable to a broader range of fidelities, scales, and applications than
other simulators, due to the flexibility and generality of its
\emph{\gls{ABM}} paradigm and \emph{discrete, object-oriented approach}.

This structure recognizes that specialists should utilize their time and
resources in modeling the specific process associated with their area of
expertise (e.g., reprocessing and advanced fuel fabrication), without having to
create a model of the entire fuel cycle to serve as its host.  \Cyclus supports
them by separating the problem of modeling a physics-dependent supply chain
into two distinct components: a simulation kernel and archetypes that interact
with the kernel. The kernel is responsible for the deployment and interaction
logic of entities in the simulation.  Physics calculations and customized
behaviors are implemented within archetypes.

Ultimately, modeling the evolution of a physics-dependent, international
nuclear fuel supply chain is a multi-scale problem which existing tools cannot
support. They have either focused on macro effects, e.g., the fleet-level
stocks and flows of commodities, or micro effects, e.g., the used-fuel
composition of fast reactor fuel. Each focus has driven the development of
specialized tools, rendering the task of answering questions between the macro
and micro levels challenging within a single tool.  In contrast, the open, extensible
architecture and discrete object tracking of \Cyclus allow the creation and
interchangeability of custom archetypes at any level of fidelity and by any
fuel cycle analyst.

\input{fcsreqs}