### 2015-03-02 fundamentals/acks.tex
Adds acknowldegements draft. Still may need some explicit legal language for NEUP, NRC etc. Acknowledges @rakhimov and @zwelchWI, which fixes #45 and fixes #43

\section{Acknowledgements}
This work has been supported by a number of people. The authors
would like to thank software contributors Zach Welch and Olzhas Rakhimov.
Additionally, the authors are grateful to the many enthusiastic members of the
\Cyclus community and our \gls{NEUP} collaboration partners at the University
of Texas, University of Utah, and University of Idaho.

This work has received institutional support from \gls{ANL}, the \gls{NEUP},
the \gls{NRC}, the \gls{NSF}, and the \gls{UW}.