### 2015-03-05 fundamentals/acks.tex
Massage the acknoledgements

\section{Acknowledgements}
This work has been supported by a number of people. The authors
would like to thank software contributors Zach Welch and Olzhas Rakhimov.
Additionally, the authors are grateful to the many enthusiastic members of the
\Cyclus community and our \gls{NEUP} collaboration partners at the University
of Texas, University of Utah, and University of Idaho.

Direct support for the \Cyclus project has been received from the \gls{NRC}
Faculty Development program, the \gls{NEUP} Research \& Development Program,
the \gls{NNSA} Consortium for Verification Technology, and the \gls{UW}.  In
addition students contributing to \Cyclus have received fellowship support
from the \gls{ANL} Lab Grad program, the \gls{NEUP} Fellowship program, the
\gls{NSF} Graduate Fellowship program, and the \gls{DHS} Nuclear Forensics
Graduate Fellowship program.