### 2015-03-11 fundamentals/abs.tex
adjusts abstract

The \Cyclus fuel cycle simulation framework and its modeling ecosystem are
presented. This tool incorporates modeling methodologies to enable flexible
simulation fidelity as well as contribution and collaboration among experts. To
date, current tools are often rigid rather than flexible, fleet-based rather
than discrete, or privately distributed rather than open source. Each of these
choices presents a challenge to modeling fidelity, generality, efficiency,
robustness, and scientific transparency.  \Cyclus incorporates modern insights
from simulation science and software architecture to solve these problems so
that challenges in nuclear fuel cycle analysis can be better addressed.
Implementation of these insights in the \Cyclus framework and its modeling
ecosystem is described. Finally, the current capabilities of \Cyclus are
demonstrated for both an open and closed fuel cycle.