### 2015-03-11 cover_letter/cover_letter.tex
adds cover letter

% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\topmargin=-1in % Moves the top of the document 1 inch above the default
\textheight=10in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
%\oddsidemargin=-10pt % Position of the left margin, can be negative or positive if you want more or less room
%\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right
%\usepackage{wallpaper}
%\ULCornerWallPaper{0.97}{bk_logo.eps}

\begin{document}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{N. Tsoulfanidis and A. Klein\\
Editors\\
Nuclear Technology}

%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\begin{center}
\includegraphics[width=\textwidth]{bk_logo.eps}
\end{center}

\address{Department of Nuclear Engineering\\
4155 Etcheverry Hall\\
MC 1730\\
University of California, Berkeley\\
Berkeley, CA 94720-1730}

\name{\vspace{5mm}
Kathryn Huff\\
Postdoctoral Scholar\\
huff@berkeley.edu}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Dear Dr. Klein and Dr. Tsoulfanidis:}

Please find enclosed a manuscript entitled: ``Fundamental Concepts in the
Cyclus Fuel Cycle Simulator Framework and Modeling Ecosystem'' which I am
submitting for exclusive consideration of publication as a research article in
Nuclear Technology.

This manuscript describes the fundamental concepts in the Cyclus fuel cycle
simulator framework and its modeling ecosystem. These concepts and their
implementations are described in the context of motivating challenges in
nuclear fuel cycle simulation. Additionally, the current capabilities of Cyclus
are demonstrated for both an open and a closed fuel cycle.

We request that this manuscript be considered for joint publication alongside
Scopatz et al., ``Cyclus Archetypes.'' These two papers are complementary, and we
believe it would be ideal for both manuscripts to be published together in the
same issue of Nuclear Technology.

Thank you for your consideration of this work. I expect it will be of interest
to a broad readership concerned with nuclear fuel cycle simulation and
computational methods. Please address all correspondence concerning this
manuscript to me at UC Berkeley and feel free to correspond with me by e-mail.

\closing{Sincere regards,%
\\
\vspace{-20mm}\includegraphics[height=1.5cm]{signature.eps}
}
%----------------------------------------------------------------------------------------

\end{letter}

\end{document}