### 2015-03-11 fundamentals/abs.tex
uses cyclus macro in abstract

\newpage
\begin{abstract}
As nuclear power expands, technical, economic, political, and environmental
analyses of nuclear fuel cycles by simulators increase in importance.  To date,
however, current tools are often rigid rather than flexible, fleet-based rather
than discrete, or privately distributed rather than open source. Each of these
choices presents a challenge to modeling fidelity, generality, efficiency,
robustness, and scientific transparency.  The \Cyclus nuclear fuel cycle
simulator framework and its modeling ecosystem incorporate modern insights from
simulation science and software architecture to solve these problems so that
challenges in nuclear fuel cycle analysis can be better addressed.  A summary
of the \Cyclus fuel cycle simulator framework and its modeling ecosystem are
presented. Additionally, the implementation of each is discussed in the context
of motivating challenges in nuclear fuel cycle simulation.  Finally, the
current capabilities of \Cyclus are demonstrated for both an open and a closed
fuel cycle.
\end{abstract}