### 2015-08-28 cover_letter/cover_letter.tex
updates cover letter for advances in engineering software

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}
 \begin{document}
\AddToShipoutPictureBG*{\includegraphics[width=\paperwidth,height=\paperheight]{background.pdf}}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Prof. A.K. Noor, Prof. B.H.V. Topping\\
Editors\\
Advances in Engineering Software}
%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\address{Kathryn D. Huff\\
4155 Etcheverry Hall\\
MC 1730\\
Berkeley, CA 94720}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Dear Professors Noor and Topping:}

Please find enclosed a manuscript entitled: ``Fundamental Concepts in the 
Cyclus Fuel Cycle Simulator Framework'' which I am submitting for exclusive 
consideration of publication as a research article in Advances in Engineering 
Software.

This manuscript describes the fundamental concepts in the Cyclus nuclear fuel 
cycle simulator framework and its modeling ecosystem. These concepts and their 
implementations are described in the context of motivating challenges in 
nuclear fuel cycle simulation. Accordingly, the current capabilities of Cyclus 
are demonstrated for a range of representative fuel cycle scenarios including 
open (no recycling), closed (complete recycling), and modified open (some 
recycling) material routing schemes.

%We request that this manuscript be considered for joint publication alongside
%Scopatz et al., ``Cyclus Archetypes.'' These two papers are complementary, and we
%believe it would be ideal for both manuscripts to be published together in the
%same issue of Advances in Engineering Software.

Thank you for your consideration of this work. I expect it will be of interest
to a broad readership concerned with nuclear engineering, industrial and 
systems engineering, fuel cycle simulation, and agent-based computational 
methods. Please address all correspondence concerning this manuscript to me at 
UC Berkeley and feel free to correspond with me by e-mail.

\closing{Sincere regards,
\includegraphics[height=2cm]{signature}\\
\fromsig{Kathryn Huff}
}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}